
\documentclass[a4paper, twoside]{report}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{array}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage[table, x11names,dvipsnames, table]{xcolor}
%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{url}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\setcounter{secnumdepth}{3}

\usepackage{fancyhdr} % page layout

\usepackage{listings}
\usepackage{color}
\usepackage{float}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}



%\setlength{\arrayrulewidth}{1mm}


\renewcommand{\footrulewidth}{0.75pt}
\renewcommand{\headrulewidth}{0.75pt}

\pagestyle{fancy}

\pagenumbering{roman}

\title{Gentoo - Webapp vulnerability detection through semi-automated black-box scanning}
\author{Abra\~{a}o Pacheco Dos Santos Peres Mota}
% Update supervisor and other title stuff in title/title.tex

\begin{document}
\input{title/title.tex}

\clearpage{\pagestyle{fancy}\cleardoublepage}
\begin{abstract}
In a world where cyber security is a growing concern in every Internet connected business and service, there is now an explosion of web application users like never before. In using these services, often times users render away valuable data about themselves - only to have it stolen in cyber attacks. It is often the case that in these attacks the vulnerability which facilitated it could have easily been fixed, but not enough emphasis is placed on web security to take preventative measures against this. \\

In the effort against this we find web application scanners, tools designed to analyse websites to detect vulnerabilities so that these may be found and removed. Many of the tools of this type out there are entirely automated, but how feasible or useful would a semi-automated, human centered tool be? \\

We develop Gentoo, a Google Chrome extension which is designed with this purpose in mind. We benchmark it against other well known web application scanners in scanning purposefully vulnerable web applications. We determine that this approach to the problem fills a particular niche that has not yet been explored by many automatic tools, leveraging the position of power that a browser extension has. We also show how Gentoo was used in finding and exploiting a real vulnerability on a live website, validating its use in a practical scenario.
\end{abstract}


\clearpage{\pagestyle{fancy}\cleardoublepage}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
	
I would like to thank Dr Sergio Maffeis for both his teaching and supervision both before and throughout this project, his help was invaluable in completing this project. \\

I want to thank Dom, Giulio, Mike and Elias for their friendship and patient help throughout my time at Imperial, and for giving me good laughs inbetween. I also want to thank Diogo for his unwavering support in the years I've known him. \\

I want to thank my brothers and sisters in Clapham, who have given me a home outside of home, and become my extended family. \\

\vspace{25mm}
Mas acima de tudo quero agradecer \`a minha familia porque sempre estiveram e sempre estar\~ao comigo. Amo-vos a todos. Sempre.
\end{abstract}

\tableofcontents
%\listoffigures
%\listoftables

\pagenumbering{arabic}
\input{introduction/introduction.tex}
\input{background/background.tex}
\input{project/project_design.tex}
\input{project/implementation.tex}
\input{project/applications.tex}
%\input{experimentation/experimentation.tex}
\input{evaluation/evaluation.tex}
%\input{future/future.tex}
\input{conclusion/conclusion.tex}
\input{appendix/appendix.tex}

\bibliographystyle{plain}
\bibliography{bibs/sample.bib}

\end{document}